Lynott et al. (2014) - Potential error in statistics reporting

### 2017-11-17 manuscript.Rmd
minor fix

In any one-sided test, for an alpha level of 5%, we can reject $H_0$ when the 90% confidence interval around the observed estimate is in the predicted direction, and does not contain the value tested against (e.g., 0). In TOST, the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$). It is not necessary to control for multiple tests when using the TOST procedure because both tests need to be statistically significant to conclude equivalence. This means that when reporting a TOST result, authors can suffice by reporting the one-sided test with the smaller test parameter (e.g. $t$ or $r$) and thus the larger $p$-value. Statistical equivalence can be concluded when the largest of the two $p$-values is smaller than alpha.

Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze (but we recommend to justify the Type 1 error rate in original research based on substantive arguments [@Lakens2017b]. It might be easiest to think of an equivalence test as checking whether the confidence interval excludes the SESOI, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a NHST is larger than some *p*-value bound.

@@

Since this is a replication, we could decide that we are interested in whether the observed effect size is smaller than the smallest effect size that the original study could have detected. This means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$). To calculate which difference corresponds to a critical z value in the original study, we multiply critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find the original study could observe a significant effect for differences of $`r printnum(crit.diffe)`$ or more.

>>>>

In any one-sided test, for an alpha level of 5%, we can reject $H_0$ when the 90% confidence interval around the observed estimate is in the predicted direction, and does not contain the value tested against (e.g., 0). In TOST, the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$). It is not necessary to control for multiple tests when using the TOST procedure because both tests need to be statistically significant to conclude equivalence. This means that when reporting a TOST result, it suffices to reporting the one-sided test with the smaller test parameter (e.g. $t$ or $r$) and thus the larger $p$-value. Statistical equivalence can be concluded when the largest of the two $p$-values is smaller than alpha.

Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze [but we recommend to justify the Type 1 error rate in original research based on substantive arguments, @Lakens2017b]. It might be easiest to think of an equivalence test as checking whether the confidence interval excludes the SESOI, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a NHST is larger than some *p*-value bound.

@@

Since this is a replication, we could decide that we are interested in whether the observed effect size is smaller than the smallest effect size that the original study could have detected. This means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$). To calculate which difference corresponds to a critical z value in the original study, we multiply critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
